---
title: "p8105_hw3_jkl2195"
author: "Jessie Li"
date: "2023-10-08"
output: github_document
---

```{r echo=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(p8105.datasets)
```

Loads the data
```{r}
data("instacart")
instacart = 
  instacart |> 
  as_tibble()

data("brfss_smart2010")
brfss_smart2010 =
  brfss_smart2010 |>
  as_tibble()
```

# Problem 1

There are a total of `r nrow(instacart)` rows and `r ncol(instacart)` cols in this dataset. This dataset contains characterization of Instacart orders. Some characterization includes product name, aisles (ex: fresh vegetables, 	
lunch meat), department(ex: produce, deli), if the product has been ordered by the same customer before, the order which customer selects each product, the days since last order, and its product and order id. The data contains `r instacart |> select(product_id) |> distinct() |> count()` products,  `r instacart |> select(user_id, order_id) |> distinct() |> count()`, orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

There are in total of `r instacart |> select(aisle) |> distinct() |> count()` aisles with the top three ordered `r instacart |> group_by(aisle) |> summarize(n = n()) |> arrange(desc(n)) |> head(3) |> unlist() |> paste(collapse = ", ")`.


This is a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered.

```{r}
instacart |>
  count(aisle) |>
  filter(n > 10000) |>
  mutate(aisle = fct_reorder(aisle, n)) |>
  ggplot(aes(x = aisle, y = n)) +
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

This is a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits” with the number of times each item is ordered

```{r}
instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |>
  count(product_name) |>
  mutate(
    rank = min_rank(desc(n))
  ) |>
  filter(rank < 4) |>
  arrange(desc(n))
```

This is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream"))|>
  group_by(product_name, order_dow) |>
  count(order_hour_of_day) |>
  summarize(
    mean = round(mean(n),2)
  ) |>
  pivot_wider(
    names_from = order_dow,
    values_from = mean
  )
```

# Problem 2

Clean up data by clean up variable names, focus on the “Overall Health” topic, removing data entries with reponses as "Very good", and convert responses as a factor taking levels ordered from “Poor” to “Excellent”.
```{r}
brfss_smart2010 = brfss_smart2010 |>
  janitor::clean_names("snake") |>
  filter(topic %in% c("Overall Health")) |>
  rename(
    state = locationabbr,
    location = locationdesc
  ) |>
  filter(!response %in% c("Very good")) |>
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Excellent"))
  )
```


**Q: In 2002, which states were observed at 7 or more locations? What about in 2010?**
```{r}
brfss_2002_multiloc_state = brfss_smart2010 |>
  filter(year == 2002) |>
  group_by(state) |>
  distinct(geo_location) |>
  summarize(num_of_locations = n()) |>
  filter(num_of_locations >= 7) |>
  mutate(
    location_full = state.name[match(state, state.abb)]
  )|>
  pull(location_full)

brfss_2010_multiloc_state = brfss_smart2010 |>
  filter(year == 2010) |>
  group_by(state) |>
  distinct(geo_location) |>
  summarize(num_of_locations = n()) |>
  filter(num_of_locations >= 7) |>
  mutate(
    location_full = state.name[match(state, state.abb)]
  )|>
  pull(location_full)
```
In 2002, `r paste(brfss_2002_multiloc_state, collapse = ", ")` (`r length(brfss_2002_multiloc_state)` total) were observed at 7 or more locations. And in 2010, `r paste(brfss_2010_multiloc_state, collapse = ", ")` (`r length(brfss_2010_multiloc_state)` total) were observed at 7 or more locations



This is a "spaghetti" plot limited of the Excellent responses of each states that maps averages the data_value across locations within a state over the years.
```{r}
brfss_smart2010_excellent = brfss_smart2010 |>
  filter(response == "Excellent") |>
  group_by(state, year) |>
  mutate(mean = mean(data_value)) |>
  select(year, state, mean)
brfss_smart2010_excellent |>
  ggplot(aes(x = year, y = mean, group = state, color = state)) +
  geom_line() +
  ggtitle("Mean Data Value over The Years Among Excellent Responses of Each State") + 
  theme(
    legend.position = "none",
    plot.title = element_text(size = 12, face = "bold")
  ) 
```


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss_2006_data_value = brfss_smart2010 |>
  filter(year == 2006, state == "NY") |>
  select(response, data_value) |>
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  labs(
    title = "2006 Distribution of Data Value \namong Responses in NY",
    x = "Response",
    y = "Data Value"
  )

brfss_2010_data_value = brfss_smart2010 |>
  filter(year == 2010, state == "NY") |>
  select(response, data_value) |>
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  labs(
    title = "2010 Distribution of Data Value \namong Responses in NY",
    x = "Response",
    y = "Data Value"
  )

brfss_2006_data_value + brfss_2010_data_value
```

# Problem 3



