---
title: "p8105_hw3_jkl2195"
author: "Jessie Li"
date: "2023-10-08"
output: github_document
---

```{r echo=FALSE}
library(tidyverse)
library(ggridges)
library(p8105.datasets)

```

Loads the data
```{r}
data("instacart")
instacart = 
  instacart |> 
  as_tibble()
```
```{r}
instacart |> 
  group_by(aisle) |>
  summarize(n = n()) |>
  arrange(desc(n)) |>
  head(3) |>
  unlist() |>
  paste(collapse = ", ")
```

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):

There are a total of `r nrow(instacart)` rows and `r ncol(instacart)` cols in this dataset. This dataset contains characterization of Instacart orders. Some characterization includes product name, aisles (ex: fresh vegetables, 	
lunch meat), department(ex: produce, deli), if the product has been ordered by the same customer before, the order which customer selects each product, the days since last order, and its product and order id. The data contains `r instacart |> select(product_id) |> distinct() |> count()` products,  `r instacart |> select(user_id, order_id) |> distinct() |> count()`, orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

There are in total of `r instacart |> select(aisle) |> distinct() |> count()` aisles with the top three ordered `r instacart |> group_by(aisle) |> summarize(n = n()) |> arrange(desc(n)) |> head(3) |> unlist() |> paste(collapse = ", ")`.


This is a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered.

```{r}
instacart |>
  count(aisle) |>
  filter(n > 10000) |>
  mutate(aisle = fct_reorder(aisle, n)) |>
  ggplot(aes(x = aisle, y = n)) +
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

This is a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits” with the number of times each item is ordered

```{r}
instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |>
  count(product_name) |>
  mutate(
    rank = min_rank(desc(n))
  ) |>
  filter(rank < 4) |>
  arrange(desc(n))
```

This is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream"))|>
  group_by(product_name, order_dow) |>
  count(order_hour_of_day) |>
  summarize(
    mean = round(mean(n),2)
  ) |>
  pivot_wider(
    names_from = order_dow,
    values_from = mean
  )
```


