---
title: "p8105_hw3_jkl2195"
author: "Jessie Li"
date: "2023-10-08"
output: github_document
---

```{r echo=FALSE}
library(tidyverse)
library(ggridges)
library(readr)
library(patchwork)
library(p8105.datasets)
```

Loads the data
```{r}
data("instacart")
instacart = 
  instacart |> 
  as_tibble()

data("brfss_smart2010")
brfss_smart2010 =
  brfss_smart2010 |>
  as_tibble()
```

# Problem 1

There are a total of `r nrow(instacart)` rows and `r ncol(instacart)` cols in this dataset. This dataset contains characterization of Instacart orders. Some characterization includes product name, aisles (ex: fresh vegetables, 	
lunch meat), department(ex: produce, deli), if the product has been ordered by the same customer before, the order which customer selects each product, the days since last order, and its product and order id. The data contains `r instacart |> select(product_id) |> distinct() |> count()` products,  `r instacart |> select(user_id, order_id) |> distinct() |> count()`, orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

There are in total of `r instacart |> select(aisle) |> distinct() |> count()` aisles with the top three ordered `r instacart |> group_by(aisle) |> summarize(n = n()) |> arrange(desc(n)) |> head(3) |> unlist() |> paste(collapse = ", ")`.


This is a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered.

```{r}
instacart |>
  count(aisle) |>
  filter(n > 10000) |>
  mutate(aisle = fct_reorder(aisle, n)) |>
  ggplot(aes(x = aisle, y = n)) +
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

This is a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits” with the number of times each item is ordered

```{r}
instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |>
  count(product_name) |>
  mutate(
    rank = min_rank(desc(n))
  ) |>
  filter(rank < 4) |>
  arrange(desc(n))
```

This is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream"))|>
  group_by(product_name, order_dow) |>
  count(order_hour_of_day) |>
  summarize(
    mean = round(mean(n),2)
  ) |>
  pivot_wider(
    names_from = order_dow,
    values_from = mean
  )
```

# Problem 2

Clean up data by clean up variable names, focus on the “Overall Health” topic, removing data entries with reponses as "Very good", and convert responses as a factor taking levels ordered from “Poor” to “Excellent”.
```{r}
brfss_smart2010 = brfss_smart2010 |>
  janitor::clean_names("snake") |>
  filter(topic %in% c("Overall Health")) |>
  rename(
    state = locationabbr,
    location = locationdesc
  ) |>
  filter(!response %in% c("Very good")) |>
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Excellent"))
  )
```


**Q: In 2002, which states were observed at 7 or more locations? What about in 2010?**
```{r}
brfss_2002_multiloc_state = brfss_smart2010 |>
  filter(year == 2002) |>
  group_by(state) |>
  distinct(geo_location) |>
  summarize(num_of_locations = n()) |>
  filter(num_of_locations >= 7) |>
  mutate(
    location_full = state.name[match(state, state.abb)]
  )|>
  pull(location_full)

brfss_2010_multiloc_state = brfss_smart2010 |>
  filter(year == 2010) |>
  group_by(state) |>
  distinct(geo_location) |>
  summarize(num_of_locations = n()) |>
  filter(num_of_locations >= 7) |>
  mutate(
    location_full = state.name[match(state, state.abb)]
  )|>
  pull(location_full)
```
In 2002, `r paste(brfss_2002_multiloc_state, collapse = ", ")` (`r length(brfss_2002_multiloc_state)` total) were observed at 7 or more locations. And in 2010, `r paste(brfss_2010_multiloc_state, collapse = ", ")` (`r length(brfss_2010_multiloc_state)` total) were observed at 7 or more locations



This is a "spaghetti" plot limited of the Excellent responses of each states that maps averages the data_value across locations within a state over the years.
```{r}
brfss_smart2010_excellent = brfss_smart2010 |>
  filter(response == "Excellent") |>
  group_by(state, year) |>
  mutate(mean = mean(data_value)) |>
  select(year, state, mean)
brfss_smart2010_excellent |>
  ggplot(aes(x = year, y = mean, group = state, color = state)) +
  geom_line() +
  ggtitle("Mean Data Value over The Years Among Excellent Responses in Each State") + 
  theme(
    legend.position = "none",
    plot.title = element_text(size = 12, face = "bold")
  ) 
```


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss_2006_data_value = brfss_smart2010 |>
  filter(year == 2006, state == "NY") |>
  select(response, data_value) |>
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  labs(
    title = "2006 Distribution of Data Value \namong Responses in NY",
    x = "Response",
    y = "Data Value"
  ) + 
  theme(
    plot.title = element_text(size = 12, face = "bold")
  )

brfss_2010_data_value = brfss_smart2010 |>
  filter(year == 2010, state == "NY") |>
  select(response, data_value) |>
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  labs(
    title = "2010 Distribution of Data Value \namong Responses in NY",
    x = "Response",
    y = "Data Value"
  ) + 
  theme(
    plot.title = element_text(size = 12, face = "bold")
  )

brfss_2006_data_value + brfss_2010_data_value
```

# Problem 3

**Background:** Accelerometers have become an appealing alternative to self-report techniques for studying physical activity in observational studies and clinical trials, largely because of their relative objectivity. During observation periods, the devices can measure MIMS in a short period; one-minute intervals are common. Because accelerometers can be worn comfortably and unobtrusively, they produce around-the-clock observations.

```{r}
substr("Hello",4,7)
```

Reads data
```{r}
nhanes_accel = read_csv("data/nhanes_accel.csv") |>
  janitor::clean_names() 



nhanes_covar = read_csv("data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  filter(
    age >= 21,
    !is.na(sex),
    !is.na(education),
    !is.na(age)
  ) |>
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female"
    ),
    education = case_match(
      education,
      1 ~ "less than high school",
      2 ~ "high school equivalent",
      3 ~ "more than high school"
    )
  )

nhanes_merged = inner_join(nhanes_accel, nhanes_covar, by = "seqn")

```

|>
  pivot_longer(
    cols = starts_with("min"),
    names_to = "min",
    values_to = "mims_value"
  ) |>
  mutate(
    min = sub('...','',min)
  )

Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.



