---
title: "p8105_hw3_jkl2195"
author: "Jessie Li"
date: "2023-10-08"
output: github_document
---

```{r echo=FALSE}
library(tidyverse)
library(ggridges)
library(readr)
library(patchwork)
library(p8105.datasets)
```

Loads the data
```{r}
data("instacart")
instacart = 
  instacart |> 
  as_tibble()

data("brfss_smart2010")
brfss_smart2010 =
  brfss_smart2010 |>
  as_tibble()
```

# Problem 1

There are a total of `r nrow(instacart)` rows and `r ncol(instacart)` cols in this dataset. This dataset contains characterization of Instacart orders. Some characterization includes product name, aisles (ex: fresh vegetables, 	
lunch meat), department(ex: produce, deli), if the product has been ordered by the same customer before, the order which customer selects each product, the days since last order, and its product and order id. The data contains `r instacart |> select(product_id) |> distinct() |> count()` products,  `r instacart |> select(user_id, order_id) |> distinct() |> count()`, orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

There are in total of `r instacart |> select(aisle) |> distinct() |> count()` aisles with the top three ordered `r instacart |> group_by(aisle) |> summarize(n = n()) |> arrange(desc(n)) |> head(3) |> unlist() |> paste(collapse = ", ")`.


This is a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered.

```{r}
instacart |>
  count(aisle) |>
  filter(n > 10000) |>
  mutate(aisle = fct_reorder(aisle, n)) |>
  ggplot(aes(x = aisle, y = n)) +
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

This is a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits” with the number of times each item is ordered

```{r}
instacart |>
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |>
  count(product_name) |>
  mutate(
    rank = min_rank(desc(n))
  ) |>
  filter(rank < 4) |>
  arrange(desc(n))
```

This is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream"))|>
  group_by(product_name, order_dow) |>
  count(order_hour_of_day) |>
  summarize(
    mean = round(mean(n),2)
  ) |>
  pivot_wider(
    names_from = order_dow,
    values_from = mean
  )
```

# Problem 2

Clean up data by clean up variable names, focus on the “Overall Health” topic, removing data entries with reponses as "Very good", and convert responses as a factor taking levels ordered from “Poor” to “Excellent”.
```{r}
brfss_smart2010 = brfss_smart2010 |>
  janitor::clean_names("snake") |>
  filter(topic %in% c("Overall Health")) |>
  rename(
    state = locationabbr,
    location = locationdesc
  ) |>
  filter(!response %in% c("Very good")) |>
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Excellent"))
  )
```


**Q: In 2002, which states were observed at 7 or more locations? What about in 2010?**
```{r}
brfss_2002_multiloc_state = brfss_smart2010 |>
  filter(year == 2002) |>
  group_by(state) |>
  distinct(geo_location) |>
  summarize(num_of_locations = n()) |>
  filter(num_of_locations >= 7) |>
  mutate(
    location_full = state.name[match(state, state.abb)]
  )|>
  pull(location_full)

brfss_2010_multiloc_state = brfss_smart2010 |>
  filter(year == 2010) |>
  group_by(state) |>
  distinct(geo_location) |>
  summarize(num_of_locations = n()) |>
  filter(num_of_locations >= 7) |>
  mutate(
    location_full = state.name[match(state, state.abb)]
  )|>
  pull(location_full)
```
In 2002, `r paste(brfss_2002_multiloc_state, collapse = ", ")` (`r length(brfss_2002_multiloc_state)` total) were observed at 7 or more locations. And in 2010, `r paste(brfss_2010_multiloc_state, collapse = ", ")` (`r length(brfss_2010_multiloc_state)` total) were observed at 7 or more locations



This is a "spaghetti" plot limited of the Excellent responses of each states that maps averages the data_value across locations within a state over the years.
```{r}
brfss_smart2010_excellent = brfss_smart2010 |>
  filter(response == "Excellent") |>
  group_by(state, year) |>
  mutate(mean = mean(data_value)) |>
  select(year, state, mean)
brfss_smart2010_excellent |>
  ggplot(aes(x = year, y = mean, group = state, color = state)) +
  geom_line() +
  ggtitle("Mean Data Value over The Years Among Excellent Responses in Each State") + 
  theme(
    legend.position = "none",
    plot.title = element_text(size = 12, face = "bold")
  ) 
```


Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss_2006_data_value = brfss_smart2010 |>
  filter(year == 2006, state == "NY") |>
  select(response, data_value) |>
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  labs(
    title = "2006 Distribution of Data Value \namong Responses in NY",
    x = "Response",
    y = "Data Value"
  ) + 
  theme(
    plot.title = element_text(size = 12, face = "bold")
  )

brfss_2010_data_value = brfss_smart2010 |>
  filter(year == 2010, state == "NY") |>
  select(response, data_value) |>
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  labs(
    title = "2010 Distribution of Data Value \namong Responses in NY",
    x = "Response",
    y = "Data Value"
  ) + 
  theme(
    plot.title = element_text(size = 12, face = "bold")
  )

brfss_2006_data_value + brfss_2010_data_value
```

# Problem 3

**Background:** Accelerometers have become an appealing alternative to self-report techniques for studying physical activity in observational studies and clinical trials, largely because of their relative objectivity. During observation periods, the devices can measure MIMS in a short period; one-minute intervals are common. Because accelerometers can be worn comfortably and unobtrusively, they produce around-the-clock observations.

```{r}
substr("Hello",4,7)
```

Reads data
```{r}
nhanes_accel = read_csv("data/nhanes_accel.csv") |>
  janitor::clean_names() 



nhanes_covar = read_csv("data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  filter(
    age >= 21,
    !is.na(sex),
    !is.na(education),
    !is.na(age)
  ) |>
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female"
    ),
    education = case_match(
      education,
      1 ~ "less than high school",
      2 ~ "high school equivalent",
      3 ~ "more than high school"
    ), 
    education = factor(education,
      levels = c("less than high school", "high school equivalent", "more than high school")
    )
  )

nhanes_merged = right_join(nhanes_accel, nhanes_covar, by = "seqn") |>
  pivot_longer(
    cols = starts_with("min"),
    names_to = "min",
    values_to = "mims_value"
  ) |>
  mutate(
    min = sub('...','',min)
  )
```


This is a table of the number of men and women in each education category

```{r}
nhanes_merged |>
  group_by(education, sex) |>
  distinct(seqn) |>
  summarise(n = n()) |>
  pivot_wider(
    names_from = sex,
    values_from = n
  )
```
For education high school equivalent, there is a significantly higher male:female ratio among the participants. For the rest of the data, the sex ratio is pretty balanced.

This is a visualization of the age distributions for men and women in each education category.
```{r fig.height = 7}
nhanes_age_less_edu = nhanes_merged |>
  filter(education == "less than high school") |>
  group_by(education, sex, age) |>
  distinct(seqn) |>
  summarise(population = n()) |>
  mutate(sex = forcats::fct_reorder(sex, population)) |>
  ggplot(aes(x = age, y = sex, fill = sex, group = sex)) +
  geom_density_ridges(alpha = 0.5) +
  ggtitle("Age Distributions for Men and Women with \nEducation Less than High School")+ 
  theme(
    plot.title = element_text(size = 12, face = "bold")
  )

nhanes_age_mid_edu = nhanes_merged |>
  filter(education == "high school equivalent") |>
  group_by(education, sex, age) |>
  distinct(seqn) |>
  summarise(population = n()) |>
  mutate(sex = forcats::fct_reorder(sex, population)) |>
  ggplot(aes(x = age, y = sex, fill = sex, group = sex)) +
  geom_density_ridges(alpha = 0.5) +
  ggtitle("Age Distributions for Men and Women with \nEducation High School or Equivalent")+ 
  theme(
    plot.title = element_text(size = 12, face = "bold")
  )

nhanes_age_high_edu = nhanes_merged |>
  filter(education == "more than high school") |>
  group_by(education, sex, age) |>
  distinct(seqn) |>
  summarise(population = n()) |>
  mutate(sex = forcats::fct_reorder(sex, population)) |>
  ggplot(aes(x = age, y = sex, fill = sex, group = sex)) +
  geom_density_ridges(alpha = 0.5) +
  ggtitle("Age Distributions for Men and Women with \nEducation More than High School")+ 
  theme(
    plot.title = element_text(size = 12, face = "bold")
  )

nhanes_age_less_edu / nhanes_age_mid_edu / nhanes_age_high_edu
```
Participants have similar age distribution between each sex with education less than high school and more than high school. For participants with education high school or equivalent, the age distribution for male tend to have a younger concentration than female, centering around age 30 to 50 compared to 45 to 75 for female group.

This is visualization of the total activity over the day over age comparing between sex.
```{r}
nhanes_merged |>
  group_by(sex, age) |>
  summarise(total_activities = sum(mims_value)) |>
  ggplot(aes(x = age, y = total_activities, color = sex, fill = sex)) +
  geom_point() +
  geom_smooth(alpha = 0.2) + 
  labs(
    title = "Age vs Total Activity over The Day for Male and Female",
    y = "total activities"
  ) + 
  theme(
    plot.title = element_text(size = 12, face = "bold")
  ) 
```
The total activities for male participants drop by age until 50. Then the total activities steadily increase at a constant rate according to the estimation. This could be dued to the outliers for two between 65 and 70 and one at age 80. Female participants tends to boost their activity until age 30 and decreases afterwards. The activities had a small boost at age 50 - 60 and another increase after age 65. Again, this could be affected by the existing outliers. In total, female participants have a higher activity values between age 25 to 60, and 72 to 80. Male participants have a higher activities values before age 25 and between age 60 to 72.


This is a plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

```{r}
nhanes_merged |>
  group_by(sex, education, min) |>
  mutate(min = as.numeric(min)) |>
  summarise(total_activities = mean(mims_value)) |> # mean or else sum will include factor such as population
  ggplot(aes(x = min, y = total_activities, group = sex, color = sex, fill = sex)) +
  geom_point(alpha = 0.1) +
  geom_smooth(alpha = 0.2) + 
  labs(
    title = "Total Activity over The Day for Male and Female \nFor Each Education Categories",
    x = "Time of The Day (min)",
    y = "total activities"
  ) + 
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(angle=45)
  ) +
  facet_grid(~education)
```

At a glance, participants increases their activities from 250 min (~4:10 am) to 500 min (~8:20am). Groups with education high school equivalent and more than high school maintain a constant activities value between 500 min(~8:20am) to 1125 min (~6:45pm). Less than high school group reached the highest at the beginning of the time frame compared to other groups and decrease to around the same level as the other groups. This could be due to lower education group tends to have a job require more physical activities than other groups. Transportation involved during communte time may also affects this trend. Female participants have a higher physical activity value at this time frame, with the highest disparity in the more than high school group. A possible explanation is that female participants tends to get busier with children, self-development, career, etc.. Lastly, the activity decreases to range 0 to 2.5 from 1125 min (~6:45pm) to 250 min (~4:10 am), with male participants having higher physical activities after 0 min (12pm).
